Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-21 aistats.tex
Go over Frederik's feedback
And spot the error in self-concordance of the log-partition of a Gaussian

\pdfsuppresswarningpagegroup=1 %removes the warning
% This is a hack, the other way to remove is to convert all pdf into postscript or eps.

@@

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\graphicspath{{figs/}}
%\usepackage{subfigure}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}       % hyperlinks
\usepackage[dvipsnames]{xcolor}

@@

	\begin{proof}
	As $\hat \mu_n = \inv{n} \sum_i X_i^2$ 
	with $X_i \sim \cN(0, \mu_*)$, the ratio to the optimum $n \frac{\hat \mu_n}{\mu_*}$
	follows a \href{https://en.wikipedia.org/wiki/Chi-square_distribution}{Chi-square distribution} with $n$ degrees of freedom, $\chi^2(n)$. Its inverse $\frac{\mu_*}{\hat \mu_n} $ follows an \href{https://en.wikipedia.org/wiki/Inverse-chi-squared_distribution}{inverse Chi-square distribution} with expectation 
	\begin{align}
		\expect[\hat \mu_n]{\frac{\mu_*}{\hat \mu_n}  - 1}  
		= \expect[\mu_n]{n \frac{\mu_*}{n \hat \mu_n}  - 1} 
		= \expect{\frac{n}{\chi^2(n)} - 1} =
		\begin{cases}
			\frac{n}{n-2} -1 = \frac{2}{n-2} \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
		\end{cases}
	\end{align}
	There is also a closed form solution for the expected logarithm of a Chi-squared, 
	because a Chi-squared is a special case of Gamma distributions, whose sufficient statistic is $(X, \log(X))$,
	\fdk{I don't follow why this is ``because it is a Gamma with sufficient statistics $(X,\log(X))$''?}
	\begin{align}
		\expect{\log \frac{\hat \mu_n}{\mu_*}} = \psi\paren{\half[n]} - \log\paren{\half[n]}
	\end{align}
	where $\psi$ is the \href{https://en.wikipedia.org/wiki/Digamma_function}{digamma function.}
	Closed forms also exist for non-central chi-squares, see for example the note of \citep{pav2015moments}.
	Consequently the suboptimality of the MLE has a closed form solution
	\begin{align}
	\expect{\bregmanconj( \mu_*; \hat \mu_n) }
	&= \half \expect{\frac{\mu_*}{ \hat \mu_n} - 1 + \log \left(\frac{\hat \mu_n}{\mu_*} \right) },
	\\
	& =
	\begin{cases}
		\half \left ( \frac{2}{n-2} +\psi(\half[n]) - \log(\half[n]) \right )  \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
	\end{cases}
	\end{align}
	Surprisingly, we need $3$ samples or more for the expected loss to be bounded and
	the optimal value $\mu_*$ does not appear in this rate, 
	so that the convergence rate is independent of the solution.
	When the expectation is finite, we can get a more interpretable formula using known \href{https://en.wikipedia.org/wiki/Digamma_function#Inequalities}{bounds on the digamma function},
	\begin{align}
		-\inv{x} \leq \psi(x) - \log(x) \leq -\inv{2x}	\; ,
	\end{align}
	giving, for $n\geq 3$,
	\begin{align}
		\inv{n-2} - \inv{n}
		&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
		\leq \inv{n-2} - \inv{2 n},
		\\
		\iff
			\frac{2}{n(n-2)}
			&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
			\leq \inv{2n} +\frac{2}{n(n-2)},
	\end{align}
	so we get a $\Omega(n^{-2})$ lower bound and a $O(n^{-1}) + O(n^{-2})$ upper bound.
	\end{proof}

\subsection{Multivariate MLE}
In higher dimensions, $X_i \sim \cN(0, \mu_*)$, $X_i\in\real^d$, and the mean parameter $\mu_*$ is a symmetric positive definite covariance matrix of order $d$, living in dimension $p = \frac{d(d+1)}{2}$. 
Note that we switched the role of $d$ from $\dim(T(x))=d$ to $\dim(x) =d$ and $\dim(T(x))=p$.
\fdk{This $\dim$ is not introduced? I would say that $\dim(T(X))$ is $d^2$ if $T(X) = XX^\top$. Slightly confused here.}

@@

 	= \half \left ( \frac{2p}{n-d-1} + \psi_d\paren{\half[n]} - d \log\paren{\half[n]}  \right ).

@@

	&= \sum_{i=0}^{d-1} \paren{\psi \paren{\half[n - i]} - \log\paren{\half[n]} },
	\\
	&\leq \sum_{i=0}^{d-1} \paren{\log\paren{\half[n-i]} - \inv{n-i} - \log\paren{\half[n]} }
	= \sum_{i=0}^{d-1} \paren{\log\paren{1 - \frac{i}{n}} - \frac{1}{n-i}} \; .

@@

	&\leq \log(n-d) + \gamma + \inv{2(n-d) - 1} - \log(n) - \gamma - \inv{2n+1},
	\\
	&= \log\paren{1 - \frac{d}{n}}  + \frac{2(d + 1)}{(2n+1)(2(n - d) - 1)}.

@@

	+ \frac{2(d + 1)}{(2n+1)(2(n - d) - 1)},

@@

 	&\leq p \paren{ \inv{n-d-1} - \inv{2n}}  + \frac{(d + 1)}{(2n+1)(2(n - d) - 1)}, 
 	\intertext{\fdk{I don't follow the second term on the next line. Is it because it is $\leq$ rather than $=$?}}
 	&\leq p \frac{n + d +1}{2 n (n-d-1)}  + \frac{(d + 1)}{2n.2 (n - d - 1)}, \\
 	&= \frac{p}{2n} + \frac{2 p(d+1)}{2 n (n-d-1)} + \frac{p/d}{2n (n - d - 1)}, \\
 	\intertext{\fdk{Why is this an inequality and }}

@@

\begin{proof}
	The lower bound can be readily obtained by applying Jensen's inequality 
	to the convex function $x\mapsto\inv{x}$ for $x>0$.
	The upper bound requires more work.
	To start, let us plug in the definition of $\MAPm$
\alignn{
		\expect{\frac{\mu^*}{\MAPm}} 
		= \expect{\frac{\MAPt}{\nat^*}} 
		= \expect{\frac{(n_0+n) \mu^*}{n_0 \mu_0 + \sum_i X_i^2}}  
		= (n_0 + n) \expect{\inv{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i^2}{ \mu^*} }} 
		= (n_0 +n) \expect{\inv{a + \chi^2(n)}},
}
	where $\chi^2(n) = \sum_i \frac{X_i^2}{ \mu^*}$ is a chi-square random variable of degree $n$ and $a=n_0 \frac{\mu_0}{\mu^*}$. The expectation reads
\alignn{
 	\expect{\inv{a + \chi^2(n)}}  
	= \int_0^\infty \inv{a+x} \frac{x^{\frac{n}{2}-1}e^{-\half[x]} }{2^{\half[n]}\Gamma(\half[n])} dx.
}
	We reformulate this expectation with the trick
	\fdk{I don't follow from there.}
\alignn{
		&\int_1^\infty e^{-\frac{zt}{2}} dt = \frac{2}{z} e^{-\half[z]} \\
		\implies &\inv{z} = \half e^{\half[z]} \int_1^\infty e^{-\frac{zt}{2}} dt \\
		\implies & \boxed{\inv{a+x} = \half e^{\frac{a+x}{2}} \int_1^\infty e^{- \frac{(a+x)t}{2}}dt \; ,}
} 
	to get two integrals instead of one
\alignn{
		2^{\half[n]}\Gamma(\half[n]) \expect{\inv{a + \chi^2(n)}}  
		&= \int_0^\infty \frac{x^{\frac{n}{2}-1}e^{-\half[x]} }{a+x} dx \\
		&= \int_0^\infty dx\  x^{\frac{n}{2}-1}e^{-\half[x]} \half e^{\half[a+x]} \int_1^\infty  dt\  e^{-\half[(a+x)t]}  \\
		&= \half e^{\half[a]} \int_1^\infty dt \ e^{-\half[a t]}   \int_0^\infty dx \  x^{\frac{n}{2}-1} e^{-\half[x t]}  \; ,
}
	where we used Fubini to switch integrals, without further justification.
	\fdk{Fubini's theorem would be a justification here, right? Are you trying to say ``assuming we can exchange the order of integration''?}
	Now with the change of variable $x = 2\frac{y}{t}$, eg $dx = 2\frac{dy}{t}$, the inner integral becomes
	\begin{align}
		\expect{\inv{a + \chi^2(n)}}  
		&= \frac{e^{\half[a]} }{2^{\half[n]+1}\Gamma(\half[n]) }\int_1^\infty dt \ e^{-\half[a t]}   \int_0^\infty 2\frac{dy}{t} \  (2\frac{y}{t})^{\frac{n}{2}-1} e^{-y} \\
		&= \frac{e^{\half[a]} }{2\Gamma(\half[n]) } 
		\underbrace{\int_1^\infty  \frac{e^{-\half[a t]}}{t^{\frac{n}{2}}}  dt}_{E_{\half[n]}(\half[a])  } \ 
		\underbrace{\int_0^\infty  y^{\frac{n}{2}-1} e^{-y} dy}_{\Gamma(\half[n])}
	\end{align}
	so we finally get  the formula valid for all $n\geq 1$
	\begin{align}
		\boxed{\expect{\inv{a + \chi^2(n)}}  = \half e^{\frac{a}{2}} E_{\half[n]}(\half[a])}
		\label{eq:hard_expectation}
	\end{align}
	where the {generalized exponential integral function} is defined as 
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; .
	\end{align}
	Now our goal is to bound this function with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	\fdk{Was this a comment from your notes that passed its way here or do you want to go with this?}
	For $n \geq 2$, eg $k\geq 1$, we have \href{https://dlmf.nist.gov/8.19.E21}{the general bound} \citep[Eq.~8.19.21]{DLMF}
	\begin{align}
		e^x E_k(x) 
		&\leq \inv{x + k - 1} \\
		\iff \expect{\inv{a + \chi^2(n)}}
		&\leq \inv{a + n - 2} \\
		\iff \expect{\frac{\mu^*}{\MAPm}} 
		&\leq \frac{n +n_0}{a + n - 2}
	\end{align}
We are left with a special case when  $n=1$, eg $k=\half$. 
Then we can use the trivial bound 
\begin{align}
	\expect{\inv{a + X^2}} < \inv{a}
	\label{eq:trivial_bound}
\end{align}
to conclude the proof.
\end{proof}

@@

The Gaussian log-partition function and entropy are, up to constants,
\alignn{
	\logpart(\nat) &= \frac{\nat_1^2}{-4\nat_2} - \half \log(-\nat_2) \\
	\conj(\m) &= - \half \log (\mu_2 - \mu_1^2) \; ,
}
where $\nat_1\in\real, \nat_2 <0$ and $\mu_1\in\real, \mu_2>\mu_1^2$.
These functions and their domains they are represented in Figure \TODO{}.
We provide definitions of self-concordance in \cref{app:self-concordant}.

\paragraph{Entropy is Self-Concordant,}
\citet[p.177]{nesterov2003introductory} proves that logarithmic barriers for a second order region are self-concordant. Adopting their notation, $\conj$ fits into this definition with $\mA = \begin{pmatrix} 2 & 0 \\ 0 & 0 \end{pmatrix}$ and $a = \begin{pmatrix} 1 \\ 0 \end{pmatrix}$.

\paragraph{Log-partition is not Self-Concordant}
Let us fix a straight line with $\nat_1 \neq 0$, e.g. $g(\nat_2) := A(\nat_1,\nat_2)$. Then
\alignn{
g'(\nat_2) = \frac{\partial A}{\partial\nat_2}(\nat) 
&= +\frac{\nat_1^2}{4 \nat_2^2} + \frac{1}{2\nat_2} \\
g''(\nat_2) = \frac{\partial^2 A}{(\partial\nat_2)^2}(\nat) 
&= -\frac{\nat_1^2}{2 \nat_2^3} - \frac{1}{2\nat_2^2} 
=  \frac{1}{2\nat_2^2}\paren{\frac{\nat_1^2}{-\nat_2^2} - 1} \\ 
g'''(\nat_2) = \frac{\partial^3 A}{(\partial\nat_2)^3}(\nat) 
&= +\frac{3\nat_1^2}{2 \nat_2^4} + \frac{1}{\nat_2^3}
= \frac{1}{2\nat_2^3}\paren{2 - 3 \frac{\nat_1^2}{- \nat_2^2}}
}
Setting $\nat_2 = - \nat_1^2$ yields $g''(\nat_2) = 0$ and $g'''(\nat_2) \neq 0$, meaning the second derivative cannot dominate the third derivative for any choice of constant or exponent. The log-partition is neither self concordant nor pseudo-self-concordant.

\fdk{TODO: Double check the above, something doesn't sit quite right, for the following. How can we get $g'' = 0$ if $A$strictly convex?}

@@

\begin{definition}[Self-concordance]
A convex function is $F:\real^p \rightarrow \real$ is self-concordant if it is differentiable $3$ times and if for all $w, v \in\real^p$ the function $g(t) = F(w+tv)$ satisfies $\abs{g'''(t)} \leq 2 g''(t)^{\frac{3}{2}}$
\end{definition}

\begin{definition}[Pseudo Self-concordance] \citep{bach2010self} 
A convex function is $F:\real^p \rightarrow \real$ is self-concordant if it is differentiable $3$ times and if for all $w, v \in\real^p$ the function $g(t) = F(w+tv)$ satisfies $\abs{g'''(x)} \leq C \norm{v}_2 g''(x)$ for some $C\geq 0$
\end{definition}

>>>>

%\pdfsuppresswarningpagegroup=1 %removes the warning
% This is a hack, the other way to remove is to convert all pdf into postscript or eps.
% RLP: not sure what is the point of this. With my latex it creates an error, and I do not have warnings, so I comment it out for now.

@@

%\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\graphicspath{{figs/}}
%\usepackage{subfigure}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}       % hyperlinks
\usepackage[dvipsnames]{xcolor}

@@

	\begin{proof}
	As $\hat \mu_n = \inv{n} \sum_i X_i^2$ 
	with $X_i \sim \cN(0, \mu_*)$, the ratio to the optimum $n \frac{\hat \mu_n}{\mu_*}$
	follows a \href{https://en.wikipedia.org/wiki/Chi-square_distribution}{Chi-square distribution} with $n$ degrees of freedom, $\chi^2(n)$. Its inverse $\frac{\mu_*}{\hat \mu_n} $ follows an \href{https://en.wikipedia.org/wiki/Inverse-chi-squared_distribution}{inverse Chi-square distribution} with expectation 
	\begin{align}
		\expect[\hat \mu_n]{\frac{\mu_*}{\hat \mu_n}  - 1}  
		= \expect[\mu_n]{n \frac{\mu_*}{n \hat \mu_n}  - 1} 
		= \expect{\frac{n}{\chi^2(n)} - 1} =
		\begin{cases}
			\frac{n}{n-2} -1 = \frac{2}{n-2} \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
		\end{cases}
	\end{align}
	There is also a closed form solution for the expected logarithm of a Chi-squared.
	Indeed a Chi-squared is a special case of Gamma distributions, whose sufficient statistic is $(X, \log(X))$, so one can apply formula~\eqref{eq:mirror-map} on the log-partition of a Gamma. The result is
	\begin{align}
		\expect{\log \frac{\hat \mu_n}{\mu_*}} = \psi\paren{\half[n]} - \log\paren{\half[n]}
	\end{align}
	where $\psi$ is the \href{https://en.wikipedia.org/wiki/Digamma_function}{digamma function.}
	Closed forms also exist for non-central chi-squares, see for example the note of \citep{pav2015moments}.
	Consequently the suboptimality of the MLE has a closed form solution
	\begin{align}
	\expect{\bregmanconj( \mu_*; \hat \mu_n) }
	&= \half \expect{\frac{\mu_*}{ \hat \mu_n} - 1 + \log \left(\frac{\hat \mu_n}{\mu_*} \right) },
	\\
	& =
	\begin{cases}
		\half \left ( \frac{2}{n-2} +\psi(\half[n]) - \log(\half[n]) \right )  \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
	\end{cases}
	\end{align}
	Surprisingly, we need $3$ samples or more for the expected loss to be bounded and
	the optimal value $\mu_*$ does not appear in this rate, 
	so that the convergence rate is independent of the solution.
	When the expectation is finite, we can get a more interpretable formula using known \href{https://en.wikipedia.org/wiki/Digamma_function#Inequalities}{bounds on the digamma function},
	\begin{align}
		-\inv{x} \leq \psi(x) - \log(x) \leq -\inv{2x}	\; ,
	\end{align}
	giving, for $n\geq 3$,
	\begin{align}
		\inv{n-2} - \inv{n}
		&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
		\leq \inv{n-2} - \inv{2 n},
		\\
		\iff
			\frac{2}{n(n-2)}
			&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
			\leq \inv{2n} +\frac{2}{n(n-2)},
	\end{align}
	so we get a $\Omega(n^{-2})$ lower bound and a $O(n^{-1}) + O(n^{-2})$ upper bound.
	\end{proof}

\subsection{Multivariate MLE}
In higher dimensions, $X \sim \cN(0, \mu_*)$, $X\in\real^d$, $T(X) = XX^\top$, and the mean parameter $\mu_*$ is a symmetric positive definite covariance matrix of order $d$, living in dimension $p = \frac{d(d+1)}{2}$. 
Note that we switched the role of $d$ from $\dim(T(\cX))=d$ to $\dim(\cX) =d$ and $\dim(T(\cX))=p$, where $\dim(\cZ)$ is the dimension of the smallest affine subspace containing $\cZ$.

@@

 	= \frac{p}{n-d-1} + \half \left ( \psi_d\paren{\half[n]} - d \log\paren{\half[n]}  \right ).

@@

	&= \sum_{i=0}^{d-1} \paren{\psi \paren{\half[n - i]} - \log\paren{\half[n]} } \\
	&\leq \sum_{i=0}^{d-1} \paren{\log\paren{\half[n-i]} - \inv{n-i} - \log\paren{\half[n]} } \\
	&= \sum_{i=0}^{d-1} \paren{\log\paren{1 - \frac{i}{n}} - \frac{1}{n-i}} \; .

@@

	&\leq \log(n-d) + \gamma + \inv{2(n-d) - 1} - \log(n) - \gamma - \inv{2n+1} \\
	&= \log\paren{1 - \frac{d}{n}}  + \frac{2(d + 1)}{(2n+1)(2(n - d) - 1)} \\
	&< \log\paren{1 - \frac{d}{n}}  + \frac{d + 1}{2n(n - d - 1)} \; .

@@

	+  \frac{d + 1}{2n(n - d - 1)} \; ,

@@

 	&\leq p \paren{ \inv{n-d-1} - \inv{2n}}  + \frac{d + 1}{4n(n - d - 1)}\\
 	&< p \frac{n + d +1}{2 n (n-d-1)}  + \frac{d(d + 1)/2}{2dn(n - d - 1)}\\
 	&= \frac{p}{2n}\frac{n-d-1}{n-d-1} + p\frac{2(d+1)}{2 n (n-d-1)} + \frac{p}{2dn (n - d - 1)}\\
 	&= \frac{p}{2n} + \frac{p(d+ 1 + \inv{2d})}{n (n-d-1)} \\

@@

\begin{proof}
	The lower bound can be readily obtained by applying Jensen's inequality 
	to the convex function $x\mapsto\inv{x}$ for $x>0$.
	The upper bound requires more work.
	To start, let us plug in the definition of $\MAPm$
\alignn{
		\expect{\frac{\mu^*}{\MAPm}} 
		= \expect{\frac{\MAPt}{\nat^*}} 
		= \expect{\frac{(n_0+n) \mu^*}{n_0 \mu_0 + \sum_i X_i^2}}  
		= (n_0 + n) \expect{\inv{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i^2}{ \mu^*} }} 
		= (n_0 +n) \expect{\inv{a + \chi^2(n)}} \; ,
}
	where $\chi^2(n) = \sum_i \frac{X_i^2}{ \mu^*}$ is a chi-square random variable of degree $n$ and $a=n_0 \frac{\mu_0}{\mu^*}$. The expectation reads
\alignn{
 	\expect{\inv{a + \chi^2(n)}}  
	= \int_0^\infty \inv{a+x} \frac{x^{\frac{n}{2}-1}e^{-\half[x]} }{2^{\half[n]}\Gamma(\half[n])} dx \; .
}
	This $1/(a+x)$ is hard to deal with.
	Fortunately, there is a trick to express it as an integral itself , with exponential terms that fit well with the $\chi^2$-density,
	\fdk{I don't follow from there.} \rlp{do you have more detailed feedback ? I don't know how to improve it.}
\alignn{
		&\int_1^\infty e^{-\frac{zt}{2}} dt = \frac{2}{z} e^{-\half[z]} \\
		\implies &\inv{z} = \half e^{\half[z]} \int_1^\infty e^{-\frac{zt}{2}} dt \\
		\implies & \boxed{\inv{a+x} = \half e^{\frac{a+x}{2}} \int_1^\infty e^{- \frac{(a+x)t}{2}}dt  \; .}
} 
We can then plug this integral into the expectation and apply Fubini's theorem to get a tractable integral on the inside
\alignn{
		2^{\half[n]}\Gamma(\half[n]) \expect{\inv{a + \chi^2(n)}}  
		&= \int_0^\infty \frac{x^{\frac{n}{2}-1}e^{-\half[x]} }{a+x} dx \\
		&= \int_0^\infty dx\  x^{\frac{n}{2}-1}e^{-\half[x]} \half e^{\half[a+x]} \int_1^\infty  dt\  e^{-\half[(a+x)t]}  \\
		&= \half e^{\half[a]} \int_1^\infty dt \ e^{-\half[a t]}   \int_0^\infty dx \  x^{\frac{n}{2}-1} e^{-\half[x t]}  \; ,
}
	assuming we can exchange the order of integration.
	%\fdk{Fubini's theorem would be a justification here, right? Are you trying to say ``assuming we can exchange the order of integration''?}\rlp{yes exactly.}
	Now with the change of variable $x = 2\frac{y}{t}$, eg $dx = 2\frac{dy}{t}$, the inner integral becomes
	\begin{align}
		\expect{\inv{a + \chi^2(n)}}  
		&= \frac{e^{\half[a]} }{2^{\half[n]+1}\Gamma(\half[n]) }\int_1^\infty dt \ e^{-\half[a t]}   \int_0^\infty 2\frac{dy}{t} \  (2\frac{y}{t})^{\frac{n}{2}-1} e^{-y} \\
		&= \frac{e^{\half[a]} }{2\Gamma(\half[n]) } 
		\underbrace{\int_1^\infty  \frac{e^{-\half[a t]}}{t^{\frac{n}{2}}}  dt}_{E_{\half[n]}(\half[a])  } \ 
		\underbrace{\int_0^\infty  y^{\frac{n}{2}-1} e^{-y} dy}_{\Gamma(\half[n])}
	\end{align}
	so we finally get  the formula valid for all $n\geq 1$
	\begin{align}
		\boxed{\expect{\inv{a + \chi^2(n)}}  = \half e^{\frac{a}{2}} E_{\half[n]}(\half[a])}
		\label{eq:hard_expectation}
	\end{align}
	where the {generalized exponential integral function} is defined as 
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; .
	\end{align}
	Now our goal is to bound this function with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	\fdk{Was this a comment from your notes that passed its way here or do you want to go with this?} \rlp{Yes I like this sentence.}
	For $n \geq 2$, eg $k\geq 1$, we have \href{https://dlmf.nist.gov/8.19.E21}{the general bound} \citep[Eq.~8.19.21]{DLMF}
	\begin{align}
		e^x E_k(x) 
		&\leq \inv{x + k - 1} \\
		\iff \expect{\inv{a + \chi^2(n)}}
		&\leq \inv{a + n - 2} \\
		\iff \expect{\frac{\mu^*}{\MAPm}} 
		&\leq \frac{n +n_0}{a + n - 2}
	\end{align}
We are left with a special case when  $n=1$, eg $k=\half$. 
Then we can use the trivial bound 
\begin{align}
	\expect{\inv{a + X^2}} < \inv{a}
	\label{eq:trivial_bound}
\end{align}
to conclude the proof.
\end{proof}

@@

The Gaussian log-partition function and entropy are, up to constants,
\alignn{
	\logpart(\nat) &= \frac{\nat_1^2}{-4\nat_2} - \half \log(-\nat_2) \\
	\conj(\m) &= - \half \log (\mu_2 - \mu_1^2) \; ,
}
where $\nat_1\in\real, \nat_2 <0$ and $\mu_1\in\real, \mu_2>\mu_1^2$.
These functions and their domains are represented in Figure \TODO{}.
We provide definitions of self-concordance in \cref{app:self-concordant}.

\paragraph{Entropy is Self-Concordant,}
\citet[Example 4.1.1.4,   p.177]{nesterov2003introductory} proves that logarithmic barriers for a second order region are self-concordant. Adopting their notation, $\conj$ fits into this definition with $\mA = \begin{pmatrix} 2 & 0 \\ 0 & 0 \end{pmatrix}$ and $a = ( 1 \ 0 )^\top$.

\paragraph{Log-partition is Self-Concordant}
Adopting the notation of definition~\ref{def:self-concordance}, let us set $w=(\nat_1 \ 0)^\top$ and $v=(0 \ -1)^\top$.
Then 
\alignn{g(t) := 2A(w+tv) = \frac{a}{t} - \log(t)}
is well defined for $t>0$ and $a=\frac{\nat_1^2}{2} \geq 0$.
The logarithm is self-concordant for the same reason that the entropy is, but the inverse function $\inv{t}$ is not \citep[example 4.1.1.5]{nesterov2003introductory}.
Linear combinations with positive coefficients  preserves self-concordance \citep[theorem 4.1.1]{nesterov2003introductory} -- e.g. self-concordance is a cone, so it seems unlikely that this sum of a self-concordant function and a not self-concordant function would be self-concordant. 
But let us verify it to be sure. The derivatives of $g$ are
\alignn{
g'(t) &= -\frac{a}{t^2} - \frac{1}{t} \; ,\\
g''(t) &= \frac{2a}{t^3} + \inv{t^2} = \frac{1}{t^2}\paren{\frac{2a}{t} + 1} \; , \\
g'''(t) &= -\frac{6a}{t^4} - \frac{2}{t^3} = \frac{-2}{t^3}\paren{\frac{3a}{t} + 1} \;.
}
Then the definition of self-concordance reads
\alignn{
\abs{g'''(t)} &\leq 2 g''(t)^{\frac{3}{2}} \\
\iff \frac{3a}{t} + 1 &\leq \paren{\frac{2a}{t} + 1}^{\frac{3}{2}} \; .
}
Writing $y = \frac{a}{t} \geq 0$, and taking the square of this expression yields
\alignn{
 (3y + 1)^2 &\leq (2y + 1)^3 \\
 \iff 9 y ^2 + 6y + 1 &\leq 8 y^3  + 12 y^2 + 6y +1 \\
\iff 0 &\leq y^2(8y + 3)
}
which is true since $y\geq 0$.

The log-partition is neither self concordant nor pseudo-self-concordant.

@@

\begin{definition}[Self-concordance]
\label{def:self-concordance}
A convex function is $F:\real^p \rightarrow \real$ is self-concordant if it is differentiable $3$ times and if for all $w, v \in\real^p$ the function $g(t) = F(w+tv)$ satisfies 
\alignn{\abs{g'''(t)} \leq 2 g''(t)^{\frac{3}{2}}}
for all feasible $t$.
\end{definition}

\begin{definition}[Pseudo Self-concordance] \citep{bach2010self} 
A convex function is $F:\real^p \rightarrow \real$ is self-concordant if it is differentiable $3$ times and if for all $w, v \in\real^p$ the function $g(t) = F(w+tv)$ satisfies 
\alignn{\abs{g'''(x)} \leq C \norm{v}_2 g''(x)}
for all feasible $t$ and for some $C\geq 0$.
\end{definition}
